Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming

### 2014-07-09 manuscripts/supplement.Rmd
fix supplement title, drop abstract

---
layout: 12pt,review
linenumbers: true
title: "Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Marc Mangel
    affiliation: cstar
  - name: Stephan Munch
    affiliation: noaa
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: marc
    address: |
      Center for Stock Assessment Research, Department of Applied Math
      and Statistics, University of California,
      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
      Biology, University of Bergen, Bergen, Norway 9020
  - code: noaa
    address: | 
      Southwest Fisheries Science Center, 
      National Oceanic and Atmospheric Administration, 
      110 Shaffer Road, Santa Cruz, CA 95060, USA

abstract: |
  Model uncertainty and limited data are fundamental challenges to
  robust management of human intervention in a natural system.
  These challenges are acutely highlighted by concerns that
  many ecological systems may contain tipping points, such as
  Allee population sizes.  Before a collapse, we do not know
  where the tipping points lie, if they exist at all.  Hence,
  we know neither a complete model of the system dynamics nor
  do we have access to data in some large region of state-space
  where such a tipping point might exist.  We illustrate how
  a Bayesian Non-Parametric (BNP) approach using a Gaussian
  Process (GP) prior provides a flexible representation of this
  inherent uncertainty. We embed GPs in a Stochastic Dynamic
  Programming (SDP) framework in order to make robust management
  predictions with both model uncertainty and limited data. We
  use simulations to evaluate this approach as compared with the
  standard approach of using model selection to choose from a set
  of candidate models.  We find that model selection erroneously
  favors models without tipping points -- leading to harvest
  policies that guarantee extinction. The GPDP performs nearly as
  well as the true model and significantly outperforms standard
  approaches. We illustrate this using examples of simulated
  single-species dynamics, where the standard model selection
  approach should be most effective, and find that it still
  fails to account for uncertainty appropriately and leads to
  population crashes, while management based on the GPDP does
  not, since it does not underestimate the uncertainty outside
  of the observed data.

keywords:
  - Bayesian
  - Structural Uncertainty
  - Nonparametric 
  - Optimal Control
  - Decision Theory
  - Gaussian Processes
  - Fisheries Management

>>>>

---
layout: 12pt,review
linenumbers: true
title: "Supplement for: Avoiding tipping points in fisheries management through Gaussian Process Dynamic Programming"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig(at)gmail.com
    footnote: Corresponding author
  - name: Marc Mangel
    affiliation: cstar
  - name: Stephan Munch
    affiliation: noaa
address: 
  - code: cstar
    address: | 
      Center for Stock Assessment Research, 
      Department of Applied Math and Statistics, 
      University of California, Mail Stop SOE-2,
      Santa Cruz, CA 95064, USA
  - code: marc
    address: |
      Center for Stock Assessment Research, Department of Applied Math
      and Statistics, University of California,
      Mail Stop SOE-2, Santa Cruz, CA 95064, USA and Department of
      Biology, University of Bergen, Bergen, Norway 9020
  - code: noaa
    address: | 
      Southwest Fisheries Science Center, 
      National Oceanic and Atmospheric Administration, 
      110 Shaffer Road, Santa Cruz, CA 95060, USA
